A Comparative Ecophysiological Survey of Lycophytes, Ferns, and Seed Plants

### 2019-02-15 manuscript/manuscript.Rmd
stomatal stats and methods written

	The goal of the current study is to examine the comparative physiology of a number of temperate ferns and lycophytes to angiosperms. We specifically examined several physiological parameters in a pairwise comparison between a fern or a lycophyte and an angiosperm in identical habitats. Given recent suggestions on inefficient stomatal functions, we hypothesized that ferns would have lower photosynthetic rates and stomatal conductance values in comparison to the nearby angiosperms. This, in turn, would impact aspects of leaf morphology (stomatal density), physiology (chlorophyll concentration), nutrient relations (N and P), and carbon relations. We speculate that the differences in physiological function may affect the ecology of these ferns, lycophytes, and angiosperms. 
#Methods

@@

This study was conducted at two sites in upstate New York. The data for most species were gathered at the Edmund Niles Huyck Preserve and Biological Research Station in Rensselaerville, New York.  This site is characterized by mixed hardwood hemlock forests along the Helderberg escarpment in the foothills of the Catskill mountain range. Data from the second site were gathered at a natural area on the campus of Colgate University in Hamilton, New York, situated in the foothills surrounding the Chenango Valley. Forests at this site are younger, but consist of a similar species richness and composition as the first site.

@@

We examined a series of fern, lycophyte, and angiosperm species at both sites (Table 1). The main focus of this work was to examine ferns relative to light-analogous angiosperms. To accomplish this, we would select a fern species and then locate the nearest neighbor angiosperm growing in similar habitat conditions. We limited our choice of angiosperms to those taxa that were herbaceous and similar in size and stature to the fern in question. At the Huyck Preserve, we sampled across two distinct habitats: understory low-light areas and open, high-light areas.

@@

To generate comparative physiological data, we followed two procedures. The first was to generate a series of detailed light response curves for a select number of species. The second approach, to increase our comparative sample size, was to make several survey measurements of maximum photosynthetic and conductance rates for additional species; this step was only completed at the Huyck Preserve. All gas exchange parameters were generated using the LiCor 6400 XT Photosynthesis System (Lincoln, NE). For the light response curves, the system was set up to a flow rate of 300 units, and a CO~2~ concentration of 400 umol m^-2 s^-1. Initial light measurements were made using the built-in quantum sensor to determine typical ambient light levels. These measurements were used to generate light response curves using appropriate light levels for high-light and low-light species separately. The light levels were not identical for each species, with several points being added or subtracted with each run, but they generally followed the same template. For high light species this was light levels of 1500, 1000, 800, 600, 500, 400, 300, 250, 200, 150, 100, 50, 25, 15, 10,5 and 0 (umol m-2 s-1). Low light curves followed this pattern; 500, 400, 350, 300, 250, 200, 150, 100, 50, 25, 15, 10, 5 and 0 (again umol m-2 s-1). Using the data we generated from the light response curves, we developed an understanding of the levels at which low-light and high-light species reached their maximum photosynthetic rates. These data were then used to complete our light saturated gas exchange survey measurements. All high-light species reached maximum rates at 1200 umol m-2 s-1 and all low-light species reached maximum rates at 600. All cuvette parameters were identical to those used for the light response curves.

@@

In open canopy habitats, ferns had significantly lower stomatal densities than those growing in closed canopies, whereas angiosperms showed no significant differences in stomatal densities between the two habitats. In both open and closed environments, ferns had significantly lower stomatal densities than analogous angiosperms (Figure asdfs). Similarly, ferns displayed a significantly lower SPAD index, a measurement of chlorophyll concentration, in open canopy environments versus closed canopies. Angiosperms showed no significant difference across the light gradient. In open environments, ferns had significantly lower SPAD index values than corresponding angiosperms, while the reverse was true in closed canopy habitats, with ferns displaying significantly higher values than analogous angiosperms (Figure asdfkj).

>>>>

	The goal of the current study is to examine the comparative physiology of a number of temperate ferns and lycophytes to angiosperms. We specifically examined several physiological parameters in a pairwise comparison between a fern or a lycophyte and an angiosperm in identical habitats. Given recent suggestions on inefficient stomatal functions, we hypothesized that ferns would have lower photosynthetic rates and stomatal conductance values in comparison to the nearby angiosperms. This, in turn, would impact aspects of leaf morphology (stomatal density), physiology (chlorophyll concentration), nutrient relations (N and P), and carbon relations. We speculate that the differences in physiological function may affect the ecology of these ferns, lycophytes, and angiosperms.

#Methods

@@

This study was conducted at two sites in upstate New York. Data for most species were gathered at the Edmund Niles Huyck Preserve (ENHP), characterized by mixed hardwood hemlock forests along the Helderberg escarpment in the foothills of the Catskill mountain range. Data from the second site were gathered at a natural area on the campus of Colgate University in Hamilton, New York, situated in the foothills surrounding the Chenango Valley. Forests at this site are younger, but consist of a similar species richness and composition at ENHP.

@@

We examined a series of fern, lycophyte, and angiosperm species at both sites (Table 1). The main focus of this work was to examine ferns relative to light-analogous angiosperms. To accomplish this, we selected a fern species and then located the nearest neighboring angiosperm growing in similar habitat conditions. Surveyed angiosperm taxa were always herbaceous and similar in size and stature to the neighboring fern species. At the ENHP, we sampled across two distinct habitats: understory low-light areas and open, high-light areas. At the Colgate University site ......

@@

To generate comparative physiological data, we followed two procedures. First, we generated a series of detailed light response curves for a select number of species from each plant group (fern = , angiosperm = ). Next, to increase our comparative sample size, we surveyed light saturated rates of photosynthesis (A~n~) and stomatal conductance (g~s~) for additional species at the ENHP. All gas exchange parameters were generated using the LiCor 6400 XT Photosynthesis System (Lincoln, NE).

For the light response curves, The CO~2~ in the leaf cuvette was set to a flow rate of 300 $\mu$mol s^-1^ and at ambient atmospheric [CO~2~] (400 ppm). Initial light measurements were made using the built-in quantum sensor to determine typical ambient light levels. These measurements were used to generate light response curves using appropriate light levels for high-light and low-light species separately. The light levels were not identical for each species, with several points being added or subtracted with each run. Broadly, light response curves followed the same template for high light or low light species. For high light species, light levels started at a PPFD of 1500 $\mu$mol m^-1^ s^-1^ and then consisted of ~16 additional steps to O $\mu$mol m^-1^ s^-1^. For low light species, light response curves started at a PPFD of 500 $\mu$mol m^-1^ s^-1^, followed by 13 additional steps to O $\mu$mol m^-1^ s^-1^. From these light response curves, we developed an understanding of the levels at which low-light and high-light species reached light saturated photosynthetic rates. As a result, additional survey measurments were conducted at at a PPFD of 1200 $\mu$mol m^-1^ s^-1^ for high-light species and 600 $\mu$mol m^-1^ s^-1^ for low-light species. All other cuvette parameters were identical to those used for the light response curves.

temperature?

@@

Overall, no differences where detected in stomatal densities between species growing open canopy and closed canopy habitats. Specifically, differences in stomatal densities were not detected for either fern nor angiosperms species between the two habitats. However, in both open and closed environments, ferns had significantly lower stomatal densities than analogous angiosperms () (Figure asdfs). In

Similarly, ferns displayed a significantly lower SPAD index, a measurement of chlorophyll concentration, in open canopy environments versus closed canopies. Angiosperms showed no significant difference across the light gradient. In open environments, ferns had significantly lower SPAD index values than corresponding angiosperms, while the reverse was true in closed canopy habitats, with ferns displaying significantly higher values than analogous angiosperms (Figure asdfkj).
